Maximum Entropy Spectral Analysis: a case study

### 2021-03-26 paper/mesa.tex
Various changes

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
\begin{abstract}
Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series. The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) or order $p$.
A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
We provide a ready to use implementation of the algorithm in the form of a python package `memspectrum`. We characterize our code performing a power spectral density analysis on some syntethic generated data (with known power spectral density) and we compare different criteria for stopping the recursion. Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational waves data from the LIGO-Virgo collaboration.
We find that, when compared to the Welch's method, the Burg's method provides a power spectral density (PSD) estimation with a systematically lower variance. This is particularly manifest in the case of a little number of datapoints, making Burg method most suitable to work in this regime.
%Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
	\title{Maximum Entropy Spectral Analysis: a case study}
	\author{Alessandro \surname{Martini}}
		\email{martini.alessandr@gmail.com}
        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
	\author{Stefano \surname{Schmidt}}
		\email{s.schmidt@uu.nl}
        \affiliation{add me}      
	\author{Walter \surname{Del Pozzo}}
		\email{walter.delpozzo@unipi.it}
        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}

@@

The study of the properties of  stochastic processes is a crucial task in many fields of physics and mathematical physics [Here we should make some examples...].
A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes. These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation. If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function 
\begin{equation}
	C(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
\end{equation}
or equivalently by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
\begin{equation}\label{eq:psd-autocorrelation}
	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
\end{equation}
In some literature, the PSD is introduced as  
\begin{equation}\label{eq:psd-f-definition}
	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
\end{equation}
without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.  
%Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time. More formally, the properties of a zero mean time series $x_t$ are described by the function:
%\begin{equation}
%	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
%\end{equation}
%where the average is computed on different times $t$.
%Thus, a stationary noise time series is completely characterized by the autocorellation function.
%\par
%More often, the study is performed in frequency domain: indeed, for stationary noise the components in frequency $\tilde{x}(f)$ of a stationary time series are conditionally independent from each other. It can be shown that:
%\begin{equation}
%	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
%\end{equation}
%This equation defines the function \textit{power spectral density} (PSD) $S(f)$.
%The PSD has a straightforward interpretation: it measures how much signal ``power" is located in each frequency bin. Furthermore, it is not difficult to show that it is the Fourier transform of the autocorrelation function:
%\begin{equation}
%	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}
%\end{equation}
%A power spectral density fully describes a stationary noisy time series and this shows how important is in the study of noise.
%\par
%Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.

An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the sample autocorrelation $\rho$, written as
\begin{equation}
    \rho_{Windowed} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
\end{equation}
where $M$ is the maximum time lag at which the autocorrelation is computed.
The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.

@@

Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
\begin{equation}\label{eq:entropy}
    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
\end{equation}
The equation can be interpreted as the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
In the continuous case:
\begin{equation}
    H[p(x)] = - \int p(x)\ln p(x) dx,
\end{equation}
We call the function $H$ (information) entropy.

@@

%
%

%This interpretation is straightforward: if we compare it with Eq.~(\eqref{eq:information}), we see that the entropy is defined as the 'average information'. The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum. Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \eqref{eq:entropy} is all we need to assign probabilities uniquely. Evidence can be added in form of constraints in the maximization.

%It is easy to verify that this expression for entropy verifies the aforementioned properties. 
%Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
%\begin{equation}
%    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
%\end{equation}
%\wdp{I need to review and correct the following}
%To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, 
%we use a simple, pedagogical example. Consider the case in which we want assign a probability distribution $p(x)$ to a variable $x$ 
%for which we know mean $\mu$ and covariance $\Sigma$. Our `evidence' is then coded as the following constraints:
%\begin{align}
%\sum_{i = 1}^{N}p_i &= 1 \\
%\sum_{i = 1}^{N}x_i p_i &= \mu \\
%\sum_{i = 1}^{N}(x_i - \mu)(x_j - \mu) p_i &= \Sigma\,
%\end{align}
%where the first row states the requirement of exhaustivity.
%\wdp{i think we want the multivariate case since i) the gaussian case is also on wikipedia ii) we actually will need that  later on.}

%First of all, we want to show that, in case in which our only constraint is exhaustiveness
%$\sum_{i = 1}^{N}p_i = 1 $,
%MAXENT reduces to indifference principle as it should be. The variational problem is defined as
%\begin{equation}\label{eq:entropyNormConst}
%    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
%\end{equation}
%with
%\begin{equation}
%    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
%\end{equation}
%so that equation \eqref{eq:entropyNormConst} reduce to
%\begin{equation}\nonumber 
%    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
%\end{equation}
%The solution of the previous equation is easily worked out requiring every term to be zero: 
%\begin{equation}\nonumber 
%    p_i = e^{-\lambda - 1}.
%\end{equation}
%Applying the normalization condition, it is immediate to show that
%\begin{equation}
%    p_i = \frac{1}{N}.
%\end{equation}
%MAXENT reduces to the IP when only the normalization constraint is used.
%
%In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
%The variational problem can be rewritten as: 
%\begin{equation}
%    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
%\end{equation}
%with constraints: 
%\begin{equation*}
%    \int \d{x} \; p(x)  = 1 \text{  and  }
%    \int \d{x} \; x p(x) = \bar x.
%\end{equation*}
%For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
%
%The general solution for $p(x)$ is:
%\begin{equation}
%    p(x) = e^{-1 - \lambda - \mu x} 
%\end{equation}
%where $\lambda$ and $\mu$ are two Lagrange multipliers
%The normalization conditions implies 
%\begin{equation}\nonumber 
%   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
%\end{equation}
%so that: 
%\begin{equation}\nonumber 
%    e^{-1 -\lambda} = \mu.
%\end{equation}
%The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
%\begin{equation}
%	\mu = \frac{1}{\bar{x}}
%\end{equation}
%The solution is then an exponential distribution:
%\begin{equation}
%    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
%\end{equation}

%Lastly, and more interestingly, we also consider a constraint on the variance. For this reason, we set:
%\begin{align*}
%    \int \d{x} \; p(x)  = 1 \text{  and  }
%	\int \d{x} \; x p(x) = \bar{x} \\
%	\text{  and  }
%	\int \d{x} \; (x-\bar{x})^2 p(x)  = \sigma^2
%\end{align*}
%Performing the same computation as above, we obtain a well known result:
%\begin{equation}
%	p(x) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\bar{x}}{\sigma}\right)^2}
%\end{equation}
%
%Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
%\begin{equation*}
%    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
%\end{equation*}
%We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals. If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}: 
%\begin{align}
%    p&\left(\vec x(t)\vert I\right) = \nonumber \\
%    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
%\end{align}

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only a few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
\par
As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
\par
Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA} \sschmidt{Which form has this filter?}. 
The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~\eqref{eq:AR_p}: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

%\sschmidt{I don't understand the next of this section...}
%Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
%Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
%Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary. So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result. \\
%Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e. it minimizes error, and is equivalent to a least square fit for the coefficients. \cite{VDBos}  \\ 
%To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
%\begin{equation}
%    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
%\end{equation}
%Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
%\begin{equation}
%E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
%\end{equation}
%which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
%Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
%\begin{equation}\nonumber 
%    \sum_{i,j}b^*_jR(j-i)b_i = P_N
%\end{equation}
%is obtained . \\ 
%The (positive) quantity  
%\begin{equation}\label{eq:linearPredictor}
%    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
%\end{equation}
%using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
%\begin{align}
%    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
%    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
%    & c^*_j R(j - i) c_i - P_N.
%\end{align}
%Moving $P_N$ on the left hand side, it becomes 
%\begin{equation}\label{eq:bestLinearPredictor}
%    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
%\end{equation}
%Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$. Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined. The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:

@@

\subsection{Choice of the loss function: LIGO Spectrum} \label{sec:LIGO_validation}
We continue our characterisation of the various loss functions considered in this work, by investigating the 
reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\wdp{add ref}.
For this analysis, we generate $500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.

@@

\sschmidt{Shall we add more?}

\section{Comparison with Welch method}\label{sec:Welch_comparison}

It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other PSD).

@@

\begin{figure}
	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{synthetic} LIGO data. The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$. In dashed black is reported the true PSD (from which data are generated).}
	\label{fig:mem_welch_fakedata}
	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_True.pdf}
\end{figure}

\begin{figure}
	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{real} LIGO data. The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
	\label{fig:mem_welch_realdata}
	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_False.pdf}
\end{figure}

@@

\sschmidt{
Some remarks:
\begin{itemize}
	\item What happens if we downsample the data? FPE should be able to capture the yearly trend...
	\item Remove yearly trend from the data?
	\item Other plots?
\end{itemize}
}

@@

\sschmidt{Other comments below??}

>>>>

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
\begin{abstract}
Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series. The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) of order $p$.
A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
We provide a ready to use implementation of the algorithm in the form of a python package `memspectrum`. We characterize our code performing a power spectral density analysis on some syntethic generated data (with known power spectral density) and we compare different criteria for stopping the recursion. Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational waves data from the LIGO-Virgo collaboration.
We find that, when compared to the Welch's method, the Burg's method provides a power spectral density (PSD) estimation with a systematically lower variance. This is particularly manifest in the case of a little number of datapoints, making Burg method most suitable to work in this regime.
%Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
	\title{Maximum Entropy Spectral Analysis: a case study}
	\author{Alessandro \surname{Martini}}
		\email{martini.alessandr@gmail.com}
        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
	\author{Stefano \surname{Schmidt}}
		\email{s.schmidt@uu.nl}
        \affiliation{Institute for Gravitational and Subatomic Physics (GRASP), Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands}      
	\author{Walter \surname{Del Pozzo}}
		\email{walter.delpozzo@unipi.it}
        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}

@@

The study of the properties of  stochastic processes is a crucial task in many fields of physics and mathematical physics \sschmidt{Here we should make some examples... Or at least cite something}.
A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes. These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation. If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function 
\begin{equation}
	C(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
\end{equation}
or equivalently by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
\begin{equation}\label{eq:psd-autocorrelation}
	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
\end{equation}
In some literature, the PSD is introduced as  
\begin{equation}\label{eq:psd-f-definition}
	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
\end{equation}
without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.  
%Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time. More formally, the properties of a zero mean time series $x_t$ are described by the function:
%\begin{equation}
%	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
%\end{equation}
%where the average is computed on different times $t$.
%Thus, a stationary noise time series is completely characterized by the autocorellation function.
%\par
%More often, the study is performed in frequency domain: indeed, for stationary noise the components in frequency $\tilde{x}(f)$ of a stationary time series are conditionally independent from each other. It can be shown that:
%\begin{equation}
%	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
%\end{equation}
%This equation defines the function \textit{power spectral density} (PSD) $S(f)$.
%The PSD has a straightforward interpretation: it measures how much signal ``power" is located in each frequency bin. Furthermore, it is not difficult to show that it is the Fourier transform of the autocorrelation function:
%\begin{equation}
%	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}
%\end{equation}
%A power spectral density fully describes a stationary noisy time series and this shows how important is in the study of noise.
%\par
%Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.

An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
\begin{equation}
    \rho_{W} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
\end{equation}
where $\rho$ is the empirical autocorrelation and $M$ is the maximum time lag at which the autocorrelation is computed.
The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.

@@

Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
\begin{equation}\label{eq:entropy}
    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
\end{equation}
The equation can be interpreted as the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
In the continuous case:
\begin{equation} \label{eq:entropy_continuos}
    H[p(x)] = - \int p(x)\ln p(x) dx,
\end{equation}
We call the function $H$ (information) entropy.

@@

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only a few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.

As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).

Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
not on the PSD.
It can be proved  \cite{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of filtering a white noise process with a filter with transfer function $S(f)$\footnote{
A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
$$T(f) = \frac{\tilde{y}(f)}{\tilde{x}(f)}$$
where $\tilde{x}(f)$ denotes the Fourier transform of $x_t$ (and similarly for $y_t$)
}.
The difference in entropy between the input and the output time series (i.e. the entropy gain) obtained by such filter applied on white noise is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}
where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
Thus maximising eq.~\eqref{eq:EntropyGain} is equivalent to maximising eq.~\eqref{eq:entropy_continuos}.

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~\eqref{eq:AR_p}: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.

Eq.~\eqref{eq:ARspectrum} and Eq.~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined. The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:

@@

\subsection{Choice of the loss function: LIGO Spectrum} \label{sec:LIGO_validation}
We continue our characterisation of the various loss functions considered in this work, by investigating the 
reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\cite{LIGO_PSDs}.
For this analysis, we generate $500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.

@@

\section{Comparison with Welch method}\label{sec:Welch_comparison}
\sschmidt{Make a good plot here...}
It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other PSD).

@@

\begin{figure}
	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{synthetic} LIGO data. The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$. In dashed black is reported the true PSD (from which data are generated).}
	\label{fig:mem_welch_fakedata}
	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
\end{figure}

\begin{figure}
	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{real} LIGO data. The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
	\label{fig:mem_welch_realdata}
	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
\end{figure}

@@

@@
